Simulating a flock of birds

### 2017-11-02 tex/article.tex
Minor clarifications
so far finished article

\usepackage{subfig}
\usepackage{todonotes}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[labelfont={bf,up},textfont={it,up}]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

@@

\begin{figure}[!htb]
	\centering
	\subfloat{{\includegraphics[width=\columnwidth/2]{low-eta-low-rho}}}
  \subfloat{{\includegraphics[width=\columnwidth/2]{low-eta-high-rho}}}
	\qquad
	\subfloat{{\includegraphics[width=\columnwidth/2]{high-eta-low-rho}}}
	\subfloat{{\includegraphics[width=\columnwidth/2]{high-eta-high-rho}}}
	\caption{Different configurations of $\eta$ and $\rho$\\Top left: low $\eta$ low $\rho$\\Top right: Low $\eta$, high $\rho$\\Bottm left: High $\eta$, low $\rho$\\Bottom right: High $\eta$, high $\rho$}\label{fig:configs}
\end{figure}

Figure~\ref{fig:configs} shows the configuration of birds for different values of $\rho$ and $\eta$. $\theta_{cone}$ is kept at $\pi$.
For low values of both $\rho$ and $\eta$ the birds tend to form smaller groups in which the birds move in the same direction.
The groups themselves move randomly.
When the noise is high and the density low, the overall motion of the birds is just random.
In the case of high noise and high density, the birds appear to move randomly but with an overall trend in directions
shared among the all the birds.
If the density is high enough, while the noise has low values, the birds move orderly in one direction.
This overall trend in motion is similar to the system going through a kinetic phase transition
where all the particles share the same direction. In order to characterize the behaviour of the flock
we have determined the average normalized velocity of the system and investigated how it changes when the input parameters are modified:

@@

To determine the critical noise parameter $\eta_c$ we used a quadratic fit model
to to fit the falling curve for $\eta < \eta_c$ and a linear fit for $\eta > 
\eta_c$, both visible in figure~\ref{fig:va_over_eta}. $\eta_c$ is then set to 
the intersection of the fit curves. Hereby we determine $\eta_c = 4.25$

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\columnwidth]{va_over_eta}
  \caption{Dependance of alignment on randomness}\label{fig:va_over_eta}
\end{figure}

@@

The cone-like vision has little effect on the final alignment of the birds.
The effect only becomes apparent near the critical value of $\rho=0.1$ after which
the alignment increases dramatically as shown in Figure~\ref{fig:va_over_rho}.
In Figure~\ref{fig:conevision} we see how only the configuration with $\rho$ near the
transition between the unordered and ordered states is affected by the implementation of the vision.
There is as expected a slight upward trend in alignment as the angle of the vision is increased.
A much more interesting result is the shape of the formed groups. In stead of forming
dense groups, the birds now form lines or open triangles whose shapes are defined
by the angle of vision. An example of such a configuration is seen in Figure~\ref{fig:angle-birds}.

@@

The shape of the formed groups when implementing the cone-like vision was a surprising result.
The triangular shape is commonly seen in migratory birds but is usually ascribed to other factors.
Our simpel model of simply letting the birds do as other birds within their sight reproduces this
effect. It is easily explained by considering two birds. One approaches the other from the side.
The first bird will see the other and follow it, whereas the second bird will see nothing and just
continue. In this way the birds form a line. If birds approach from the other side, they will line
up in a similar fashion causing a combined triangular shape. Once a traingle has formed it is much
more likely for a new approaching bird to align with one of the sides of the trangle rather than
joining the group somewhere else. In this way the triangles keep growing.
That the effect on alignment is only detectable in a limited range also makes sense, because for
low densities, it will be unlikely for birds to come close enough to each other to form groups.
For higher densities there will more often be a bird within any other bird's vision. This causes
larger, coherent groups to form.

@@

In this paper we have presented a model for studying the motion of a flock of birds.
The model contained contained few variables; density of birds, noise in their direction and an angle
controlling which other birds influence a given bird's motion. Each of these variables gave rise
to the movement of the flock going between disordered and ordered. For the case of the angle based
vision the effects on alignment was only detectable near the transition between the unordered
and ordered state for the density of birds.
However the vision based approach gave rise to different shapes of the groups formed when the density
was sufficiently low. When the vision approach was not used, the groups were dense and clumped
together. When the vision was implemented, the birds formed lines or triangular groups.

>>>>

\usepackage{subcaption}
\usepackage{todonotes}
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage[english]{babel} % Language hyphenation and typographical rules
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[labelfont={bf,up},textfont={it,up}]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

@@

\begin{figure}[!htb]
	\centering
        \begin{subfigure}[b]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{low-eta-low-rho}
            \caption{Low $\eta$, low $\rho$}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
            \includegraphics[width=\columnwidth]{low-eta-high-rho}
            \caption{Low $\eta$, high $\rho$}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{high-eta-low-rho}
            \caption{High $\eta$, low $\rho$}
        \end{subfigure}
        \begin{subfigure}[b]{0.45\columnwidth}
	    \includegraphics[width=\columnwidth]{high-eta-high-rho}
            \caption{High $\eta$, high $\rho$}
        \end{subfigure}
	\caption{Different configurations of $\eta$ and $\rho$}\label{fig:configs}
\end{figure}

Figure~\ref{fig:configs} shows the configuration of birds for different values of $\rho$ and $\eta$. $\theta_{cone}$ is kept at $\pi$.
For low values of both $\rho$ and $\eta$ the birds tend to form smaller groups in which the birds move in the same direction.
The groups themselves move randomly.
When the noise is high and the density low, the overall motion of the birds is just random.
In the case of high noise and high density, the birds appear to move randomly but with an overall trend in directions
shared among all the birds.
If the density is high enough, while the noise has low values, the birds move orderly in one direction.
This overall trend in motion is similar to the system going through a kinetic phase transition
where all the particles share the same direction. In order to characterize the behaviour of the flock
we have determined the average normalized velocity of the system and investigated how it changes when the input parameters are modified:

@@

To determine the critical noise parameter $\eta_c$ we used a quadratic fit model
to to fit the falling curve for $\eta < \eta_c$ and a linear fit for $\eta > 
\eta_c$, both visible in figure~\ref{fig:va_over_eta}. $\eta_c$ is then set to 
the intersection of the fit curves. Hereby we determine $\eta_c = 4.25$ for $N=400$.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=\columnwidth]{va_over_eta}
  \caption{Dependance of alignment on randomness for $\rho = 4$}\label{fig:va_over_eta}
\end{figure}

@@

The cone-like vision has little effect on the final alignment of the birds.
The effect only becomes apparent near the critical value of $\rho=0.1$ after which
the alignment increases dramatically as shown in Figure~\ref{fig:va_over_rho}.
In Figure~\ref{fig:conevision} we see how only the configuration with $\rho$ near the
transition between the unordered and ordered states is affected by the implementation of the vision.
There is, as expected, a slight upward trend in alignment as the angle of the vision is increased.
A much more interesting result is the shape of the formed groups. In stead of forming
dense groups, the birds now form lines or open triangles whose shapes are defined
by the angle of vision. An example of such a configuration is seen in Figure~\ref{fig:angle-birds}.

@@

The shape of the formed groups when implementing the cone-like vision was a surprising result.
The triangular shape is commonly seen in migratory birds but is usually ascribed to other factors.
Our simple model of simply letting the birds do as other birds within their sight reproduces this
effect. It is easily explained by considering two birds. One approaches the other from the side.
The foremost bird will see nothing and just continue whereas the approaching bird will see the
other and follow it.
In this way the birds form a line. If birds approach from the other side, they will line
up in a similar fashion causing a combined triangular shape. Once a triangle has formed it is much
more likely for a new approaching bird to align with one of the sides of the trangle rather than
joining the group somewhere else. In this way the triangles keep growing.
That the effect on alignment is only detectable in a limited range also makes sense, because for
low densities, it will be unlikely for birds to come close enough to each other to form groups.
For higher densities there will more often be a bird within any other bird's vision. This causes
larger, coherent groups to form.

@@

In this paper we have presented a model for studying the motion of a flock of birds.
The model contained few variables: density of birds, noise in their direction and an angle
controlling which other birds influence a given bird's motion. Each of these variables gave rise
to the movement of the flock going between disordered and ordered. For the case of the angle based
vision the effects on alignment was only detectable near the transition between the unordered
and ordered state for the density of birds.
However the vision based approach gave rise to different shapes of the groups formed when the density
was sufficiently low. When the vision approach was not used, the groups were dense and clumped
together. When the vision was implemented, the birds formed lines or triangular groups.
